Splitting of the homology of the punctured mapping class group

### 2020-05-01 referee report/Bianchi_answer_referee.tex

\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[a4article]{geometry}

\newcommand{\N}{\mathbb{N}}    
\newcommand{\C}{\mathbb{C}}
% \newcommand{\cC}{\mathcal{C}}
% \newcommand{\D}{\mathbb{D}}
% \newcommand{\F}{\mathbb{F}}
% \newcommand{\cF}{\mathcal{F}}
% \newcommand{\cK}{\mathcal{K}}
% \newcommand{\ff}{\mathfrak{f}}
\newcommand{\fM}{\mathfrak{M}}
% \newcommand{\Q}{\mathbb{Q}}
% \newcommand{\R}{\mathbb{R}}
% \renewcommand{\H}{\mathcal{H}}
% \newcommand{\fH}{\mathfrak{H}}
% \newcommand{\V}{\mathcal{V}}
% \newcommand{\Z}{\mathbb{Z}}
% \newcommand{\cS}{\mathcal{S}}
% \newcommand{\bS}{\mathbb{S}}
\newcommand{\fS}{\mathfrak{S}}
% \newcommand{\T}{\mathcal{T}}

% \newcommand{\pa}[1]{\left(#1\right)}
% \newcommand{\abs}[1]{\left|#1\right|}
% \newcommand{\set}[1]{\left\{#1\right\}}

\renewcommand{\phi}{\varphi}

\hyphenation{sta-bi-li-ty}

\begin{document}
\title{
Answer to the referee report for ``Splitting of the homology of the
punctured mapping class group''
}
\author{Andrea Bianchi}
\address{Mathematics Institute,
University of Bonn,
Endenicher Allee 60, Bonn,
Germany
}
\email{bianchi@math.uni-bonn.de}
\date{\today}
\maketitle

First, I would like to thank the referee for the meticulous work he/she made
in reviewing my article.

I have accepted almost all corrections and suggested improvements, but in the following cases
I have followed a different way than the one proposed by the referee; I will describe briefly
the reasons for my choice, and I remain open for further discussion.
\begin{itemize}
 \item The referee writes \emph{p.3, line 2: The condition that $g\geq 2$ is needed for $\mathrm{Diff}(\Sigma_{g,1})$ to be contractible.}\\
 Since we are working with surfaces with non-empty boundary, this result should also be true in genera $0$ and $1$; differently
 the components of $\mathrm{Diff}^+(\Sigma_g)$ are not contractible for $g=0,1$. I have briefly reminded the reader of this fact in the article.
 \item The referee writes \emph{p.19, line −11: ``$\mathcal{N}' = (\mathcal{N} \setminus c_1 ) \cup \mathcal{D}$''
 Since $\mathcal{N}$ is a neighborhood of $\mathcal{D} \cup c_1 \cup\dots\cup c_m$, it contains the neighborhood of
 $c_1$. It seems that the author wants to take the complement of the neighborhood
 of $c_1$ in the definition of $\mathcal{N}'$.}\\
 This is not necessary: $\mathcal{N}' = (\mathcal{N} \setminus c_1 ) \cup \mathcal{D}$ is already an open set, contains
 $\mathcal{D} \cup c_2 \cup\dots\cup c_m$ and is disjoint from $c_1\setminus\mathcal{D}$. One can reduce further
 $\mathcal{N}'$ by deleting an entire neighborhood of $c_1$, and then taking again the union with $\mathcal{D}$, but
 this is not necessary for the argument to work. I have stressed in the article that $\mathcal{N}' = (\mathcal{N} \setminus c_1 ) \cup \mathcal{D}$
 is already open.
 \item The referee writes \emph{Section 6: How about mentioning the articles ``Stable cohomology of the mapping
 class group with symplectic coefficients and the universal Abel-Jacobi map'' written by Eduard Looijenga
 and ``Characteristic classes of surface bundles'' written
 by Shigeyuki Morita. They compute the integral stable cohomology ring of the
 mapping class group of punctured surface with boundary.}\\
 It seems to me that Morita's article (BULLETIN   (New  Series)  OF  THE  AMERICAN  MATHEMATICAL   SOCIETY
 Volume  11,  Number  2, pages 386-388,  October   1984) only deals with characteristic classes of
 $\Sigma_g$-bundles (i.e., no puncture, no boundary) over $\mathbb{Q}$. Looijenga's article
 (Journal of Algebraic Geometry. 5. January 1994) also deals mainly with coefficients in $\mathbb{Q}$.
 It seems to me that the only with coefficients in $\mathbb{Z}$ in Looijenga's article is Proposition 2.1:
 however, only the case of \emph{ordered} punctures is treated, whereas the case of unordered
 punctures follows easily only over $\mathbb{Q}$, but, in my opinion, not over $\mathbb{Z}_2$.
 The stable ($g=\infty$), unordered ($m<\infty$ unordered punctures) case with coefficients
 in $\mathbb{Z}_2$ (actually any field) is rather treated in the article of B\"odigheimer and Tillmann,
 which I already cite. Therefore I would leave these two suggested references out. 
\end{itemize}
\end{document}

### 2020-05-01 arxiv version/Actiongg.tex
Final version ready to appear in JTop

\subsection{Bigradings of the form $(m,m)$}
In this subsection we construct a $\gg$-~equivariant isomorphism 
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]
% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
% both by Poincaré-Lefschetz duality.

We already know that these $\Z_2$-vector spaces have the same dimension:
indeed $\Sym_m(\H)$ is precisely the summand in bigrading $(m,m)$ in equation
\eqref{eq:isovectorspaces}, using that a monomial
 whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$.
% i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
% is $m$ it must lie in $\Sym_m(\H)$.
% 
% The rough idea in constructing the map $\psi_m$ is the following: $\Sym_m(\H)$ is generated
% by elements of the form $[c_1]\cdot\ldots\cdot[c_m]$, where $c_1,\dots,c_m\colon\Sone\to\mrS$ are simple
% closed curves; we consider the product map 
% \[
% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m(\mrS)
% \]
% which carries the fundamental class of the $m$-fold torus to some class in
% $H_m(\SP^m(\mrS))$; we then lifts this class, in a nice way, to a class
% in $H_m(\cms)$.

@@

\begin{defn}
\label{defn:cCm}
Let $\cC^m$ be the the (discrete) set of isotopy classes of $m$-tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
such that
%\begin{itemize}
%  \item each $c_r\colon\Sone\to\mrS$ is smooth and transverse to each $\U_i$ and $\V_i$;
% \item
 any two curves $c_i,c_j$ intersect each other only inside $\D$.

% \item each curve $c_i$ is transverse to $\partial\D$.
%\end{itemize}
Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
images. Here and in the following $\Sone$ is the unit circle in $\C$.

@@

An element of $\cC^m$ is called \emph{multicurve}; by abuse of notation the $m$-tuple
$(c_1,\dots,c_m)$ will often represent its class in $\cC^m$, i.e. the corresponding multicurve.
See picture \ref{fig:defcCm}
% For each isotopy class we
% suppose to have fixed a representative $(c_1,\dots,c_m)$ made of curves that are
% also transverse to all $\U_i$'s and $\V_i$'s. We write
% $[c_1,\dots,c_m]$ for the isotopy class whose preferred representative is
% $(c_1,\dots,c_m)$.

@@

\begin{figure}\centering
 \includegraphics[scale=0.7]{figures/defcCm.pdf}
 \caption{A multicurve in $\cC^5$.}
\label{fig:defcCm}
\end{figure}

% Our aim is to define a $\gg$-equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
% factors through $\pr$ to give the desired $\gg$-equivariant map $\psi$; we will then show that
% $\psi$ is an isomorphism.

% The map $\tpsi_m$ that we construct will have two additional properties. The first
% is the following: for any
% $(c_1,\dots,c_m)$ representing a class in $\cC^m$, the homology
% class $\tpsi_m(c_1,\dots,c_m)$ can be represented by a singular cycle in $\cms$ which
% is \emph{supported} on the union $\D\cup c_1\cup \dots\cup c_m$, meaning that
% this singular cycle only hits configurations in $\cms$ whose points actually
% lie in $\D\cup c_1\cup \dots\cup c_m$.
% 
% For the second property we need a definition
\begin{defn}
 \label{defn:cmsD}
 Recall Definition \ref{defn:SP}.
 The space $\cmsD$ is the subspace of $\SP^m(\mrS)$ of configurations where all points
 having multiplicity $\geq 2$ lie inside $D$.
 See Figure \ref{fig:defcmsD}.

@@

Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^{m+1}$.
% and suppose that we have proved the existence of the lift $\tpsi_m\colon \ZcC{m}\to H_m(\cms)$
% with the aforementioned property.

\begin{defn}
\label{defn:variationsCm}
We introduce several variations of the notion of configuration space; see Figure \ref{fig:defcmsD}.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{white} (meaning that
 it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
 from each other.
%  \item The space $\fmsD$ is the subspace of $\mrS\times\cmsD$ containing all configurations
%  $(q;\set{p_1,\dots,p_m})$ where either $q\in\D$ may coincide with some $p_i$ (which can appear
%  with multipliciy $\geq 1$), or $q\not\in\D$ must be distinct from all $p_i$'s.
%  Again $q$ is thewhite point
 \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots, P_m}}$ where either $\bar P\in\D$ may coincide with \emph{exactly}
 one $P_i$, or
 $\bar P\not\in \D$ must be distinct from all $P_i$'s.
 Again $\bar P$ is called the white point.
 %We see that $C_{1,m}\subset\fmstwoD$ is an open subspace.
 \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}(\mrS)$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
 somewhere in $\mrS$, with multiplicity $1$.
\end{itemize}
 \end{defn}

\begin{figure}\centering
 \includegraphics{figures/defcmsD.pdf}
 \caption{A configuration in each of the space introduced in Definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
 Whenever a multiplicity is not specified, it is equal to 1.}
\label{fig:defcmsD}
\end{figure}

@@

There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the white point
into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
% the latter being even a $m+1$-fold covering map;
we have a commutative diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \fmstwoD %@>\subset >> \fmsD
\\   @V\p VV @V\p VV %@V\p VV
\\   C_{m+1}(\S) @>\subset >> \cmstwoD %@>\subset >> C_{m+1}^D(\S)
   \end{CD}
\end{equation}

@@

\subsection{General bigradings $(m-l,m)$.} Fix $0\leq l\leq m$ for the whole subsection: our next aim is to prove Theorem
\ref{thm:Hbms*as*ggrep} for the bigrading $(m-l,m)$.
% \begin{defn}
% %  Let $\tilde\D'$ be the open trapeziod contained in $(0,1)^2\subset\mrS$ which is delimited
% %  by the four points $(1/8;1)$, $(1/4;1/2)$, $(3/4;1/2)$ and $(7/8;1)$ (SEE PICTURE); in particular
% %  $\D\subset\D'$.
% %  Let $\S'\subset\S$ be the closure in $\S$ of $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
% %  Note that $\S'$ is also a surface of type $\sg$.
% %  
%  Recall from section \ref{Preliminaries} that $\S'$ is the closure in $\S$ of
%  $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
%  
%  For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in Definition \ref{defn:cms},
%  using $\D$ and $\mrS'$ instead of $\mrS$, respectively.
%  
%  For every splitting $m=p+(m-p)$ there is an open embedding
%  \[
%  \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
%  \]
%  given by taking the union of configurations:
% 
% %  For the rest of the section let $\gg$ denote the group of isotopy classes of diffeomorphisms
% %  of $\S$ relative to $\partial\S\cup\D'$: in symbols $\gg=\pi_0\pa{\Diff(\S;\partial\S\cup\D'}$.
% \end{defn}
% Recall from section \ref{sec:Preliminaries} that $\S'$ is the closure in $\S$ of
% $\S\setminus\bar\D$.%, where $\bar\D$ is the closure of $\D$ in $\S$.

@@

We will prove the following lemma, from which Theorem \ref{thm:Hbms*as*ggrep} follows:
\begin{lem}
 \label{lem:oplussplitting}
For all $l\leq p\leq m$ the map $\mu_*$ in equation \eqref{eq:mu*} is injective, and the collection
of all these maps yields a splitting
\begin{equation}
  \label{eq:musplitting}
 H_{m-l}(\cms)=\bigoplus_{p=l}^m H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}.
\end{equation}
\end{lem}
\begin{proof}
Note that the statement of the lemma does not depend on the the action of $\gg$:
we have a map from the right-hand side to the left-hand side of equation \eqref{eq:musplitting},
we already know that it is $\gg$-equivariant,
we only need to show that it is a linear isomorphism. Note also that Lemma \ref{lem:gensymchain}
implies that the two vector spaces have the same dimension.
% 
% We can equivalently consider the maps in cohomology
%  \[
%    \mu^*\colon  H^{m-l}\pa{\cms}\to H^p\pa{C_{m-p}(\S')}\otimes H^{p-l}\pa{C_p(\D)},
%  \]
% and show that they are surjective for $0\leq p\leq m-l$ and together exhibit a product splitting
% \[
%   H^{m-l}(\cms)=\prod_{p=0}^{m-l} H^p\pa{C_{m-p}(\S')}\otimes H^{p-l}\pa{C_p(\D)}.
% \]
% Note that the natural target for the cohomology map $\mu^*$ is the whole group
% \[
%  H^{m-l}\pa{C_{m-p}(\S')\times C_p(\D)}\simeq \prod_{q=0}^p H^q\pa{C_{m-p}(\S')}\otimes H^{m-l-q}\pa{C_p(\D)},
% \]
% where we have used the K\"unneth formula; we consider here, by abuse of notation,
% the composition of this map with the projection
% on the factor corresponding to $q=p$.
% 
% Using Poincaré-Lefschetz duality and again the K\"unneth formula we reduce to show that the maps
% \[
% \mu^{\infty}_* \colon \tH_{m-l}(\cms^{\infty})\to \tH_p\pa{C_{m-p}(\S')^{\infty}}\otimes \tH_{p-l}\pa{C_p(\D)^{\infty}}
% \]
% are surjective and exhibit $\tH_{m-l}(\cms^{\infty})$ as a product
% \[
% \tH_{m-l}(\cms^{\infty})=\prod_{p=0}^{m-l}\tH_p\pa{C_{m-p}(\S')^{\infty}}\otimes \tH_{p-l}\pa{C_p(\D)^{\infty}}.
% \]
% Here a few explanations are required:
% we are considering the map $\mu^{\infty}\colon \cms^{\infty}\to \pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}$ which collapses the complement
% of the open subspace $C_{m-p}(\S')\times C_p(\D)\subset\cms$ to the point at infinity; it induces a map
% $\mu^{\infty}_*\colon \tH_{m-l}(\cms^{\infty}) \to \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}}$;
% applying Poincaré-Lefschetz duality
% and the K\"unneth formula to the second group we get an isomorphism
% \[
%  \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}}\simeq \prod_{q=0}^p \tH_q\pa{C_{m-p}(\S')}\otimes\tH_{m-l-q}\pa{C_p(\D)};
% \]
% we are considering the composition of the map $\mu^{\infty}_*$ with the projection onto the factor $q=p$, and by
% abuse of notation we call this map simply $\mu^{\infty}_*$.
%

@@

\begin{figure}\centering
 \includegraphics[scale=0.65]{figures/axb.pdf}
 \caption{From left to right, the class $[a]=\epsilon\cdot(Q\epsilon)^2\in H_2(C_5(\D))$;
 the class $[b]=\v_1\cdot\u_2\cdot\v_3^2\in H_4(C_4(\S'))$; and the product class
 $\mu_*([a]\otimes[b])\in H_6(C_9(\S))$.}
\label{fig:axb}
\end{figure}

@@

% $[a]\otimes [b]$ has a dual class in $\tH_{p-l}(C_p(\D)^{\infty})\otimes \tH_p(C_{m-p}(\S')^{\infty})$
% represented by the product of generalised symmetric chains
% \[
%  (\alpha_j)_j\otimes (p,(u_i,v_i)_{i\leq g}).
% \]
% If $\tup=(l,(x_i)_{i\leq l})$ is a tuple such that $e^{\tup}$ appears in the symmetric
% chain $(\alpha_j)_j$, and if $\tup'=(0,(u_i,v_i)_{i\leq g})$ is the only tuple such that
% $e^{\tup'}$ appears in the generalised symmetric chain $(p,(u_i,v_i)_{i\leq g})$, then
% the product $e^{\tup}\times e^{\tup'}\subset C_p(\D)\times C_{m-p}(\S')\subset\cms$ is
% contained in the closure of the generalised symmetric chain $e^{\tup}$, where...

% If $p<p'$ then
% the \emph{geometric} intersection between the cycle $|p',(\alpha'_j),(u'_i,v'_i)|$
% in $\cms^{\infty}$ and the cycle $\mu_*(a\otimes b)$ in $\cms$ is empty: indeed the first cycle is supported
% on $\infty\in\cms^{\infty}$ and on configurations where at least $p'$ points lie in the union
% $\bigcup_i(\U_i\cup\V_i)\subset\S'$,
% whereas the second cycle is supported on configurations where at least $m-p$ points lie
% inside $\D$. The algebraic intersection between the homology classes $\mu_*([a]\otimes [b])$
% and $[p',(\alpha'_j),(u'_i,v'_i)]$ is therefore also zero.
% 
% Consider now the case $p=p'$.

@@

% We replace the map $\mu^{\infty}$ by a cellular approximation that agrees with it on the $p$-skeleton,
% and by abuse of notation we still call $\mu^{\infty}$ the new map.
% 
Consider now the generalised symmetric chain $\kappa(p',\ualpha',\uu',\uv')$ representing
a class in $\tH_{m+l}(\cms^{\infty})=H^{m-l}(\cms)$, with $\ualpha'=(\alpha'_j)_{j\geq 0}$,
$\uu'=(u_1,\dots,u'_g)$ and $\uv'=(v'_1,\dots,v'_g)$; in particular $l=\sum_{j\geq 0}\alpha'_j$.
Suppose moreover $p'\leq p$.

@@

Then the matrix of algebraic intersections between these two sets of
classes is an upper-triangular matrix
with $1$'s on the diagonal, and in particular
it is invertible. This shows that the set of classes of the form $\mu_*([a]\otimes [b])$
is a basis for $H_{m-l}(\cms)$.

% 
% 
% 
% , and let $e^{\tup}$
% be one of the cells appearing in this generalised symmetric chain, with
% $\tup=(l,(x_i)_{i\leq l}, (u_i,v_i)_{i\leq g})$, and consider the restriction
% of $\mu^{\infty}\colon\cms^{\infty}\to \pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}$
% to $e^{\tup}\subset\cms^{\infty}$
% 
% If $p\neq p'$, the whole cell $e^{\tup}$ is mapped
% to $\infty$: indeed if $p<p'$, then a configuration in $e^{\tup}$ contains
% at least $p'$ points inside $\S'$, so it cannot lie in $C_p(\D)\times C_{m-p}(\S')$;
% if $p'<p$ then a configuration in $e^{\tup}$ contains at least $m-p$ points 
% 
% 
% then the algebraic
% intersection between $(p',(u'_i,v'_i),(\alpha'_j))$ and $\mu_*(a\otimes b)$ is $0\in\Z_2$,
% unless $p=p'$, $\alpha_j=\alpha'_j$ for all $j$, and $u_i=u'_i$ and $v_i=v'_i$ for all $1\leq i\leq g$.
% 
% Indeed the generalised symmetric chain $(p',(u'_i,v'_i),(\alpha'_j))$ hits only configurations
% in $\cms$ where for all $1\leq i\leq g$ at least $u'_i$ points lie on $\U_i$ and at least $v'_i$
% points lie on $\V_i$
\end{proof}

@@

\begin{figure}\centering
 \includegraphics[scale=0.6]{figures/counterexample.pdf}
 \caption{$\mu_*(a\otimes b)$ and $e^{\tup'}$ intersect once, transversely inside $C_2(\S)$.}
\label{fig:counterexample}
\end{figure}

>>>>

\subsection{Bigradings of the form $(m,m)$}
In this subsection we construct a $\gg$-~equivariant isomorphism 
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]

We already know that these $\Z_2$-vector spaces have the same dimension:
indeed $\Sym_m(\H)$ is precisely the summand in bigrading $(m,m)$ in equation
\eqref{eq:isovectorspaces}, using that a monomial
 whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$.

@@

\begin{defn}
\label{defn:cCm}
Let $\cC^m$ be the the (discrete) set of isotopy classes of $m$-tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
such that any two curves $c_i,c_j$ intersect each other only inside $\D$.

Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
images. Here and in the following $\Sone$ is the unit circle in $\C$.

@@

An element of $\cC^m$ is called \emph{multicurve}; by abuse of notation the $m$-tuple
$(c_1,\dots,c_m)$ will often represent its class in $\cC^m$, i.e. the corresponding multicurve.
See picture \ref{fig:defcCm}

@@

\begin{figure}\centering
 \includegraphics[scale=0.7]{defcCm.pdf}
 \caption{A multicurve in $\cC^5$.}
\label{fig:defcCm}
\end{figure}

\begin{defn}
 \label{defn:cmsD}
 Recall Definition \ref{defn:SP}.
 The space $\cmsD$ is the subspace of $\SP^m(\mrS)$ of configurations where all points
 having multiplicity $\geq 2$ lie inside $D$.
 See Figure \ref{fig:defcmsD}.

@@

Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^{m+1}$.

\begin{defn}
\label{defn:variationsCm}
We introduce several variations of the notion of configuration space; see Figure \ref{fig:defcmsD}.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{white} (meaning that
 it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
 from each other.
 \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots, P_m}}$ where either $\bar P\in\D$ may coincide with \emph{exactly}
 one $P_i$, or
 $\bar P\not\in \D$ must be distinct from all $P_i$'s.
 Again $\bar P$ is called the white point.
 \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}(\mrS)$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
 somewhere in $\mrS$, with multiplicity $1$.
\end{itemize}
 \end{defn}

\begin{figure}\centering
 \includegraphics{defcmsD.pdf}
 \caption{A configuration in each of the space introduced in Definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
 Whenever a multiplicity is not specified, it is equal to 1.}
\label{fig:defcmsD}
\end{figure}

@@

There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the white point
into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
we have a commutative diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \fmstwoD
\\   @V\p VV @V\p VV
\\   C_{m+1}(\S) @>\subset >> \cmstwoD
   \end{CD}
\end{equation}

@@

\subsection{General bigradings $(m-l,m)$.} Fix $0\leq l\leq m$ for the whole subsection: our next aim is to prove Theorem
\ref{thm:Hbms*as*ggrep} for the bigrading $(m-l,m)$.

@@

We will prove the following lemma, from which Theorem \ref{thm:Hbms*as*ggrep} follows:
\begin{lem}
 \label{lem:oplussplitting}
For all $l\leq p\leq m$ the map $\mu_*$ in equation \eqref{eq:mu*} is injective, and the collection
of all these maps yields a splitting
\begin{equation}
  \label{eq:musplitting}
 H_{m-l}(\cms)=\bigoplus_{p=l}^m H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}.
\end{equation}
\end{lem}
\begin{proof}
Note that the statement of the lemma does not depend on the the action of $\gg$:
we have a map from the right-hand side to the left-hand side of equation \eqref{eq:musplitting},
we already know that it is $\gg$-equivariant,
we only need to show that it is a linear isomorphism. Note also that Lemma \ref{lem:gensymchain}
implies that the two vector spaces have the same dimension.

@@

\begin{figure}\centering
 \includegraphics[scale=0.65]{axb.pdf}
 \caption{From left to right, the class $[a]=\epsilon\cdot(Q\epsilon)^2\in H_2(C_5(\D))$;
 the class $[b]=\v_1\cdot\u_2\cdot\v_3^2\in H_4(C_4(\S'))$; and the product class
 $\mu_*([a]\otimes[b])\in H_6(C_9(\S))$.}
\label{fig:axb}
\end{figure}

@@

@@

Consider now the generalised symmetric chain $\kappa(p',\ualpha',\uu',\uv')$ representing
a class in $\tH_{m+l}(\cms^{\infty})=H^{m-l}(\cms)$, with $\ualpha'=(\alpha'_j)_{j\geq 0}$,
$\uu'=(u_1,\dots,u'_g)$ and $\uv'=(v'_1,\dots,v'_g)$; in particular $l=\sum_{j\geq 0}\alpha'_j$.
Suppose moreover $p'\leq p$.

@@

Then the matrix of algebraic intersections between these two sets of
classes is an upper-triangular matrix
with $1$'s on the diagonal, and in particular
it is invertible. This shows that the set of classes of the form $\mu_*([a]\otimes [b])$
is a basis for $H_{m-l}(\cms)$.
\end{proof}

@@

\begin{figure}\centering
 \includegraphics[scale=0.6]{counterexample.pdf}
 \caption{$\mu_*(a\otimes b)$ and $e^{\tup'}$ intersect once, transversely inside $C_2(\S)$.}
\label{fig:counterexample}
\end{figure}
